### 2019-05-22 doc/chapters/introduction.tex
Updated chapter Stainless in introduction.tex

Stainless is a framework developed by "Lab for Automated Reasoning and Analysis" (LARA) at EPFL's School of Computer and Communication Sciences to verify Scala programs.
It explores all input values, reports inputs for which a program fails and demonstrates counterexamples which violate a given specification.
The specification must be provided by a developer through giving a precondition and postcondition for functions.

Constrains for input values are specified in a precondition with the keyword \textit{require}. 
Restrictions for received output values are defined in a postcondition with the keyword \textit{ensuring}. 
While compiling Stainless tries to prove that for the given input the postcondition of output always holds.
Three outcomes of the verification are possible: valid, unvalid and unknown when the framework is unable to prove or find a counterexample.
The code to be verified should be written in Pure Scala belonging to functional programming paradigm. To extend this subset of the Scala language Stainless supports also some imperative features translating them into Pure Scala concepts.
For more details about Stainless its official website can be explored. \cite{Stainless}

>>>>

Stainless is a framework developed by "Lab for Automated Reasoning and Analysis" (LARA) at EPFL's School of Computer and Communication Sciences. The framework is used to verify Scala programs.
It verifies statically that a program satisfies a specification given by a developer and that a program will not crash at runtime.
Stainless explores all possible input values, reports inputs for which a program fails and demonstrates counterexamples which violate a given specification.

The main functions used to write a specification are \textit{require} and \textit{ensuring}. 
A precondition should be written at the beginning of a function body with \textit{require}. Its argument is a boolean expression which corresponds to constraint for inputs of a function being verified.
A postcondition should be written after a function body with  \textit{ensuring}. It is a verification condition on an output of a function.
While compiling Stainless tries to prove that the postcondition always holds, assuming a given precondition does hold.

3 outcomes of verification with Stainless are possible: valid, unvalid and unknown. 
If the postcondition is valid, Stainless could prove that for any inputs constrained in a precondition, the postcondition always holds.
With invalid postcondition the framework could find at least one counterexample which satisfies a precondition but violates a postcondition.
Also an output unknown is possible when Stainless is unable to prove a postcondition or find a counterexample. In this case a timeout or an internal error occured.
Furthermore, it will be verified by Stainless that a precondition cannot be violated.

The following example demonstrates a simple structure of formal specification for calculating factorial.

\begin{lstlisting}[caption=Example of verifying the function calculating factorial of an integer number, captionpos=b, language=Scala]
    def factorial(n: Int): Int = {
        require(n >= 0)
        if (n == 0) {
          1
        } else {
          n * factorial(n - 1)
        }
    } ensuring(res => res >= 0)
\end{lstlisting}

The function recursively calculates factorial of an integer number. An input to the function is constrained in \textit{require} with positive value. A result of calculation should be also positive, what will be verified by Stainless.
While compiling Stainless disproves the postcondition and gives the number 17 as a counterexample. A number of type Int is a 32 bit value. That is why the function factorial of 17 causes an overflow and results to a negative value. The program would work correct by changing type of a number to BigInt.
The outcome of Stainless verification of the function factorial is shown bellow.

\begin{figure}[H]
	\centering
		\includegraphics[scale=0.5]{images/output1.png}
	\caption{Output of Stainless verification for calculating factorial of Int number}
	\label{fig:output1}
\end{figure}

The code to be verified should be written in Pure Scala belonging to functional programming paradigm. To extend this subset of the Scala language Stainless supports also some imperative features translating them into Pure Scala concepts.
For more details about Stainless its official website can be explored. \cite{Stainless}
